A Generalized Visbility to Power-Spectrum Framework for 21\,cm Inteferometric Probes of Reionizations

### 2017-07-02 Rotation paper/draft_working.tex
save

%\documentclass[preprint2,numberedappendix,tighten,twocolappendix]{aastex6}  % USE THIS TO MAKE BIB, THEN FORMAT USING EMULATEAPJ
\documentclass[twocolumn,apj,numberedappendix]{emulateapj}
\shorttitle{Adding Sensitivity to 21\,cm Inteferometric Probes of Reionization}
\shortauthors{Zhang\&Parsons}
\usepackage{float}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[figuresright]{rotating}
%\usepackage{rotating}
\usepackage{natbib}
%\usepackage{pdflscape}
%\usepackage{lscape}
\usepackage{ctable}
\citestyle{aa}
\renewcommand\[{\begin{equation}}
\renewcommand\]{\end{equation}}
\graphicspath{{../figures/}}

@@

\caption{The PAPER-128 layout. Each dot corresponds to the location of
an antenna. Top panel shows the antenna positions drawn to scale;
bottom panel show the antenna labels and distances of the 112-element grid, in other words excluding the outrigger antennas.
The numbering of the antennas in the bottom panel are original labels
during instrument assembly and does not bear significant
meaning. In the bottom panel, the two baselines (5,77) and (40,85) marked with red segments are example of an equivalent pair, both with separation denoted by \{1, 0\}, for the
antennas are separated by 1 unit east and 0 unit north. Similarly,
the baselines marked in blue and yellow are examples
of the classes \{1, 1\} and \{1,-1\}, respectively. Due to the small North-South separation within the grid, \{1, 1\} is expected to be near-equivalent to \{1, 0\}. }
\label{fig:AntPos}
\end{figure}

We shall use the 128-element PAPER array (henceforth as PAPER-128) to motivate our formalism and demonstrate our method, and extend our results to several HERA configurations in Sec \ref{sec:arrconf}. 
The PAPER array is located in the Karoo desert in South Africa (30:43:17.5
S, 21:25:41.8 E). The layout pattern with antenna labels are show
in Fig. \ref{fig:AntPos}. The array consists of a 112-element grid and 16 "outriggers" used primarily to aid calibration.  In the bottom panel, the two baselines (5, 77) and (40, 85), marked in red, are an example of equivalent pair. We denote a equivalency-class of baselines in the PAPER grid by their separations, in this case  \{1, 0\}, for the
antennas are separated by 1 unit east and 0 unit north. Similarly,
the baselines marked in blue and yellow are respectively examples
of \{1, 1\} and \{1, -1\}.
Note \{1, 0\} and \{-1, 0\} for example are the same class and should
not be counted twice. Antennas in purely north-south baselines
are close (4m), and hence these baselines are not suitable
for cross-multiplication due to cross-talks. On the other hand, the close North-South separation means that classes such as \{1, 0\} and  \{1, 1\} are expected to be near-equivalent. The PAPER-64 analysis of \cite{Ali2015} used three classes of baselines, the PAPER-128
equivalent of which are 
\{2, 0\}, \{2, 1\} and \{2, -1\}. There each of these classes
of baselines are cross multiplied within itself. This paper provides the method for inter-class multiplications.

\section{Method}

\subsection{Rough Idea}

Given a point source on the sky, each baseline maps the
source to a point in $uv$ plane. As the Earth rotates with respect to
the source, the point traces out tracks in the $uv$ plane. 
We show in Fig. \ref{fig:Tracks} $uv$ tracks of PAPER-128 over 4.8 hours, at 0.15GHz, for a source that passes through zenith.

\begin{figure}[H]
\includegraphics[width=\linewidth]{tracks128}
\caption{Tracks of PAPER-128 (grid only, excluding outriggers) for requency $\nu=0.15\text{GHz}$ and a hypothetical source that passes through zenith.
These tracks are traced out over 0.2 sidereal days, or roughly 4.8
hours. Color represents different baselines. 
As the Earth rotates, tracks are traced out counterclockwise. }
\label{fig:Tracks}
\end{figure}

Roughly speaking, we can identify
redundancy of near-equivalent baselines as crossings
of the $uv$ tracks. As we see in Fig. \ref{fig:Tracks}, there are many such crossings. Crossing tracks, however does not imply perfect redundancy. The reason lies in the finite size of the beams. We show sample beams of HERA and PAPER antennas in Fig. \ref{fig:Beam} for reference. The inaccuracy is expected to be even more significant for the smaller HERA beam, which has a larger corresponding $uv$ point spread function.

Thus although it is a valid quick method to identify some baselines to cross-multiply, track-crossing is
not accurate enough for time offset determination, nor can give estimate of the degree of redundancy. To determine the time-offset corresponding to maximal redundancy, we must develop a more general  formalism that accounts for the point spread function of the finite beams, and estimates the degree of redundancy for general combination of baselines at a general time-offset, as we do in the next section.

@@

\caption{Sample beam response of HERA (left) and PAPER (right) antennas, both
for frequency of $\nu=150\text{MHz}$ and Stokes $I$ polarization. Notice $A$ in this paper is a ``baseline's beam'', equivalent to squares of the antenna voltage beams shown here. The circles centered around zenith (center of beam) here are
spaced 10 degrees apart. The finite size of the beams limits the accuracy of the $uv$ track-crossing as a method of redundancy search. \label{fig:Beam}}
\end{figure}

\subsection{Formalism}
Below we shall derive theoretical expectations of cross multiplications
of two near-equivalent baselines. More precisely we shall relate
the product of visibilities of two different baselines to the power spectrum.

We take the visibility as commonly defined in the literature (e.g.
\citealt{first-paper}): 
\begin{equation}
\begin{aligned}V_{\nu}(\boldsymbol{b}) & =\int d\Omega A_{\nu}(\hat{\boldsymbol{s}})\phi(\nu)I_{\nu}(\hat{\boldsymbol{s}})\exp\left[-2\pi i\frac{\nu}{c}\boldsymbol{b}\cdot\hat{\boldsymbol{s}}\right],\\
 & \approx\frac{2k_{B}}{\lambda^{2}}\int d\Omega A_{\nu}(\hat{\boldsymbol{s}})\phi(\nu)T(\hat{\boldsymbol{s}})\exp\left[-2\pi i\frac{\nu}{c}\boldsymbol{b}\cdot\hat{\boldsymbol{s}}\right],
\end{aligned}
\label{eq:Vis1}
\end{equation}
Here $\lambda$ is a mean wavelength, $\boldsymbol{b}$ is the baseline
length, $\hat{\boldsymbol{s}}$ and $\Omega$ are a direction in the
sky and its corresponding solid angle. $A_{\nu}$ is the (frequency
dependent) beam, and $I$ is the specific intensity, which has been
related to $T$, the brightness temperature in the Rayleigh-Jeans
limit. $k_B$ is Boltzmann's constant. Note that $A$ here is the baseline beam, i.e. product of two antenna voltage beams, examples of which are shown in Fig. \ref{fig:Beam}. $\phi(\nu)$ is the frequency bandpass profile. In practice,
power spectrum measurements are typically taken from a few ten MHz centered around the corresponding redshift of interest (e.g. 150 MHz for z=9.5). We take a moment to note that in Eq. \ref{eq:Vis1}, no flat-sky approximation has yet been made. The angular integral is performed over the dome $d\Omega$.

We define the delay-transformed visibility \citep{delay-transform}:
\small
\begin{equation}
\begin{aligned}V(\boldsymbol{b},\tau) & =\int d\nu V_{\nu}(\boldsymbol{b})\phi(\nu)\exp\left[-2\pi i\nu\tau\right],\\
 & =\frac{2k_{B}}{\lambda^{2}}\int d\Omega d\nu A(\hat{\boldsymbol{s}},\nu)\phi(\nu)T(\hat{\boldsymbol{s}},\nu)\exp\left[-2\pi i\nu\left(\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}+\tau\right)\right]
\end{aligned}
.\label{eq:Vb1}
\end{equation}
\normalsize

Eq. \eqref{eq:Vb1} expresses the delay-transformed visibility as
an integral over observation coordinates $\hat{\boldsymbol{s}}$ and $\nu$. Ultimately,
we would like to relate the data, collected with coordinates $\hat{\boldsymbol{s}}$
and $\nu$, to the power spectrum, written with cosmological coordinates
$\boldsymbol{r}$ and $\boldsymbol{k}$. We start by noticing that
\[
\begin{aligned}r & =\frac{c}{H_{0}}\int_{0}^{z}\frac{dz'}{E(z')},\\
 & \approx\frac{c}{H_{0}}\int_{0}^{z_{0}}\frac{dz'}{E(z')}-\frac{c(1+z)^{2}}{\nu_{21}H_{0}E(z)}\left(\nu-\nu_{0}\right),\\
 & \equiv D_{c}-Y\Delta\nu,
\end{aligned}
\]
where $\nu_{21}=1420$MHz is the 21\,cm transition rest frequency, $\nu_{0}$
a reference central frequency with corresponding redshift $z_{0}$,
and 
\[
E(z)=\sqrt{\Omega_{m}(1+z)^{3}+\Omega_{\Lambda}}.
\]
Inverting for $\nu$:
\begin{equation}
\nu=\frac{D_{c}-r}{Y}+\nu_{21}.\label{eq:nur}
\end{equation}

Further we shall invoke the flat-sky limit
\begin{equation}
(r_{x},r_{y})=(X\hat{\boldsymbol{s}}_{x},X\hat{\boldsymbol{s}}_{y}), 
\end{equation}
where
\begin{equation}
X = \frac{c}{H_0}\int_{0}^{z}\frac{dz'}{E(z')}
\end{equation}
is a cosmological scalar. In the same limit $(Xk_{x},Xk_{y},Yk_{z})=\frac{2\pi}{c}(b_{x},b_{y},\tau)$ relates
the cosmological $k$ to the observation coordinates. For a more rigorous treatment without the flat sky limit, see Appendix [ADRIAN].

We can rewrite the delayed transformed visibility as 
\small
\[
\begin{aligned}V(\boldsymbol{b},\tau) & =\frac{2k_{B}}{\lambda^{2}}\int_{H}\frac{d^{3}r}{X^{2}Y}A(\boldsymbol{r})\phi(r)T(\boldsymbol{r})\exp\left[-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\hat{\boldsymbol{r}}+\tau\right)\nu_{r}\right]\end{aligned},
\]
\normalsize
 where $d\nu=-dr/Y$ and $d^{3}r=-X^{2}Yd\Omega d\nu$.

We have written $\nu_{r}$ to remind us that $\nu$ and $r$ are related
by Eq. \eqref{eq:nur}. The beam reception pattern $A$ is dimensionless,
normalized to 1 at its peak (zenith), and we assume it to be the same
for all baselines.

With a time offset, the beam pattern has moved relative to the sky.
Here we choose to fix the sky, and denote the rotated coordinates
of the beam pattern with the 3 dimensional rotation operator $\Gamma$. With implicit bounds
of integrals from $-\infty$ to $\infty$, we have:
\begin{widetext}
\begin{equation}
\begin{aligned} & \langle V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau')\rangle\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\langle T^{*}(\boldsymbol{r})T(\boldsymbol{r'})\rangle\right)A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}\frac{d^{3}\kappa'}{(2\pi)^{3}}\langle T^{*}(\boldsymbol{\kappa})T(\boldsymbol{\kappa'})\rangle e^{-i(\boldsymbol{\kappa}\cdot \boldsymbol{r}-\boldsymbol{\kappa'}\cdot\boldsymbol{r'})}\right)A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}P(\kappa)e^{-i\boldsymbol{\kappa}\cdot(\boldsymbol{r}-\boldsymbol{r'})}\right)A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
 & \approx\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}rd^{3}r'\delta_{D}^{(3)}(\boldsymbol{r}-\boldsymbol{r'})A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}r|A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r})||\phi(\nu_{r})|^{2}\exp\left[-i2\pi\nu_{r}\left(\hat{\boldsymbol{r}}\cdot\frac{\boldsymbol{b}}{c}-\Gamma \hat{\boldsymbol{r}}\cdot\frac{\boldsymbol{b'}}{c}\right)\right],\\
 & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma \hat{\boldsymbol{s}},\nu)||\phi(\nu)|^{2}\exp\left[-i2\pi\nu\left(\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b}}{c}-\Gamma\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b'}}{c}\right)\right],
\end{aligned}
\label{eq:main}
\end{equation}

where in transition from cosmological coordinates back to observing coordinates we have written $\hat{\boldsymbol{r}}\equiv\hat{\boldsymbol{s}}$, and 
\begin{equation}
\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'})=|{\phi^{*}}(\ensuremath{\nu_{r}})\phi(\nu_{r'})|\exp\left[-i\frac{2\pi}{c}\left(\boldsymbol{b}\cdot\nu_{r}\hat{\boldsymbol{r}}-\boldsymbol{b'}\cdot\nu_{r'}\Gamma\hat{r'}\right)\right]\exp\left[-i2\pi\tau\left(\nu_{r}-\nu_{r'}\right)\right].
\end{equation}
\end{widetext}
The second to third line of Eq.(\ref{eq:main}) follows from assumption of translational invariance, and the third to fourth line follows from the assumption that
the 3D power spectrum varies negligibly over the $k$-space of interest
so that $\hat{P}_{21}(k+k_{2})\approx\hat{P}_{21}(k)$. Since $\Gamma$
is a sky rotation, it does not affect $\nu$, hence we have taken $\nu_{r}$
outside the parenthesis. Notice that the phase factor $\exp\left[-i2\pi\tau\left(\nu-\nu'\right)\right]$
drops out in the end. This means that correlation is peaked at the same time-offset for all delay channels.

Finally, since the beam pattern and bandwidth are given in $\hat{\boldsymbol{s}}$
and $\nu$, we convert the integral back to these coordinates to get
the general relation between the delay-transformed visibilities and
the power spectrum:

\begin{equation}
\begin{aligned} & \langle V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau')\rangle\\
 & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu)||\phi(\nu)|^{2}\\
 & \qquad \qquad \qquad \qquad \exp\left[-i2\pi\nu\left(\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b}}{c}-\Gamma\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b'}}{c}\right)\right].\end{aligned}
\label{eq:final}
\end{equation}

>>>>

%\documentclass[preprint2,numberedappendix,tighten,twocolappendix]{aastex6}  % USE THIS TO MAKE BIB, THEN FORMAT USING EMULATEAPJ
\documentclass[twocolumn,apj,numberedappendix]{emulateapj}
\shorttitle{Adding Sensitivity to 21\,cm Inteferometric Probes of Reionization}
\shortauthors{Zhang, Liu \&Parsons}
\usepackage{float}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[figuresright]{rotating}
%\usepackage{rotating}
\usepackage{natbib}
\usepackage{gensymb}
%\usepackage{pdflscape}
%\usepackage{lscape}
\usepackage{ctable}
\citestyle{aa}
\renewcommand\[{\begin{equation}}
\renewcommand\]{\end{equation}}
\graphicspath{{../figures/}}

@@

\caption{The PAPER-128 layout. Each dot corresponds to the location of
an antenna. Top panel shows the antenna positions drawn to scale;
bottom panel show the antenna labels and distances of the 112-element grid, in other words excluding the outrigger antennas.
The numbering of the antennas in the bottom panel are original labels
during instrument assembly and does not bear significant
meaning. In the bottom panel, the two baselines (5,77) and (40,85) marked with red segments are example of an equivalent pair, both with separation denoted by \{1,0\}, for the
antennas are separated by 1 unit east and 0 unit north. Similarly,
the baselines marked in blue and yellow are examples
of the classes \{1,1\} and \{1,-1\}, respectively. Due to the small North-South separation within the grid, \{1,1\} is expected to be near-equivalent to \{1,0\}. }
\label{fig:AntPos}
\end{figure}

We shall use the 128-element PAPER array (henceforth as PAPER-128) to motivate our formalism and demonstrate our method, and extend our results to several HERA configurations in Sec \ref{sec:arrconf}. 
The PAPER array is located in the Karoo desert in South Africa (30:43:17.5
S, 21:25:41.8 E). The layout pattern with antenna labels are show
in Fig. \ref{fig:AntPos}. The array consists of a 112-element grid and 16 "outriggers" used primarily to aid calibration.  In the bottom panel, the two baselines (5, 77) and (40, 85), marked in red, are an example of equivalent pair. We denote a equivalency-class of baselines in the PAPER grid by their separations, in this case  \{1,0\}, for the
antennas are separated by 1 unit east and 0 unit north. Similarly,
the baselines marked in blue and yellow are respectively examples
of \{1,1\} and \{1, -1\}.
Note \{1,0\} and \{-1, 0\} for example are the same class and should
not be counted twice. Antennas in purely north-south baselines
are close (4m), and hence these baselines are not suitable
for cross-multiplication due to cross-talks. On the other hand, the close North-South separation means that classes such as \{1,0\} and  \{1,1\} are expected to be near-equivalent. The PAPER-64 analysis of \cite{Ali2015} used three classes of baselines, the PAPER-128
equivalent of which are 
\{2, 0\}, \{2, 1\} and \{2, -1\}. There each of these classes
of baselines are cross multiplied within itself. This paper provides the method for inter-class multiplications. To do so we will use the short hand notation $\{m,n\}:\{m',n'\}$ to denote a pair of baseline classes to be cross-multiplied.

\section{Method \label{sec:method}}\label{sec:method}

\subsection{$uvw$ tracks \label{sec:tracks}}

Given a point source on the sky, each baseline maps to a point in $uvw$ space. As the Earth rotates with respect to
the source, the points trace out tracks in the $uvw$ space. 
We show in Fig. \ref{fig:Tracks} $uvw$ tracks of PAPER-128 over 12 hours, at 0.15GHz, for a source that passes through zenith.

\begin{figure}[H]
\includegraphics[width=\linewidth]{tracks3D}
\caption{Tracks of PAPER-128 (only a randomly selected subset of 32 baselines are shown) for frequency $\nu=0.15\text{GHz}$ and a hypothetical source that passes through zenith.
These tracks are traced out over 0.5 sidereal days. Color represents different baselines. 
As the Earth rotates, tracks are traced out counterclockwise. Tracks in $uvw$ space do not cross, even though the $uv$ projections do. }
\label{fig:Tracks}
\end{figure}

Traditionally, we can identify
redundancy of near-equivalent baselines as crossings
of the $uv$ tracks, a 2 dimensional projection of Fig. \ref{fig:Tracks}. As we see in Fig. \ref{fig:Tracks}, there would be many such crossings. However, there are several reasons that $uv$ track-crossing do not imply perfect redundancy. In fact, in our case $uv$ track-crossings are
not accurate enough for time offset determination, nor can it give estimate of the degree of redundancy.  The most obvious reason is that the 3 dimensional $uvw$ tracks do not actually cross. For drift-scan arrays, even a crossing in $uvw$ space does not imply perfect redundancy. This shall become evident in the next section, after we develop a more general formalism that accounts for the point spread function of the finite beams, and estimate the degree of redundancy for general combination of baselines at a general time-offset. An explicit explanation of track-crossing will be presented in Appendix \ref{sec:appA}.

\subsection{Formalism}
In this section we formulate the relation between 
the product of visibilities from two arbitrary baselines and the power spectrum.

We begin with the visibility as commonly defined in the literature (e.g.
\citealt{first-paper}): 
\begin{equation}
\begin{aligned}V_{\nu}(\boldsymbol{b}) & =\int d\Omega A(\hat{\boldsymbol{s}},\nu)\phi(\nu)I_{\nu}(\hat{\boldsymbol{s}})\exp\left(-2\pi i\frac{\nu}{c}\boldsymbol{b}\cdot\hat{\boldsymbol{s}}\right),\\
 & \approx\frac{2k_{B}}{\lambda^{2}}\int d\Omega A(\hat{\boldsymbol{s}},\nu)\phi(\nu)T(\hat{\boldsymbol{s}})\exp\left(-2\pi i\frac{\nu}{c}\boldsymbol{b}\cdot\hat{\boldsymbol{s}}\right),
\end{aligned}
\label{eq:Vis1}
\end{equation}
Here $k_B$ is Boltzmann's constant, $\lambda$ is the mean wavelength, $\boldsymbol{b}$ is the baseline length, $\hat{\boldsymbol{s}}$ and $\Omega$ are a direction in the
sky and its corresponding solid angle. Inside the integral we have $\phi(\nu)$ the frequency bandpass profile, $A_{\nu}$ the (frequency
dependent) primary beam, and $I$ the specific intensity, which has been
related to $T$, the brightness temperature in the Rayleigh-Jeans
limit. The beam reception pattern $A$ is dimensionless,
normalized to 1 at its peak (zenith), and we assume it to be the same
for all baselines. We show sample beams of HERA and PAPER antennas in Fig. \ref{fig:Beam} for reference.

@@

\caption{Example beam response of HERA (left) and PAPER (right) antennas. Both beams are Stokes $I$ polarization antenna voltage beams, at frequency of $\nu=150\text{MHz}$ and normalized to $1$ at zenith. Notice $A$ in this paper is a ``baseline's beam'', equivalent to squares of the antenna voltage beams shown here. The circles centered around zenith (center of beam) here are
spaced 10 degrees apart. \label{fig:Beam}}
\end{figure}

In practice,
power spectrum measurements are typically taken from a few ten MHz centered around the corresponding redshift of interest (e.g. 150 MHz for z=9.5).

Delayed-transformed visibility has been widely adopted in recent years since it has been shown that foregrounds are isolated in delay space. We define the delay-transformed visibility by Fourier transforming visibility along the frequency axis\citep{delay-transform} as:
\small
\begin{equation}
\begin{aligned}V(\boldsymbol{b},\tau) & =\int d\nu V_{\nu}(\boldsymbol{b})\phi(\nu)\exp\left(-2\pi i\nu\tau\right),\\
 & =\frac{2k_{B}}{\lambda^{2}}\int d\Omega d\nu A(\hat{\boldsymbol{s}},\nu)T(\hat{\boldsymbol{s}},\nu)\exp\left[-2\pi i\nu\left(\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}+\tau\right)\right]. 
\end{aligned}
\label{eq:Vb1}
\end{equation}
\normalsize
Here the delay $\tau$ is the Fourier variable. Eq. \eqref{eq:Vb1} expresses the delay-transformed visibility as
an integral over observational coordinates $\hat{\boldsymbol{s}}$ and $\nu$. For notational simplicity here and in the rest of this paper, we shall absorb the bandpass $\phi$ inside the primary beam $A$.

Ultimately,
we would like to relate the data, collected with coordinates $\hat{\boldsymbol{s}}$
and $\nu$, to the power spectrum, written with 
$\boldsymbol{r}$ and $\boldsymbol{k}$, the cosmological position coordinate and wavenumber. To do so, we start by noticing that
\[
\begin{aligned}r & =\frac{c}{H_{0}}\int_{0}^{z}\frac{dz'}{E(z')},\\
 & \approx\frac{c}{H_{0}}\int_{0}^{z_{0}}\frac{dz'}{E(z')}-\frac{c(1+z)^{2}}{\nu_{21}H_{0}E(z)}\left(\nu-\nu_{0}\right),\\
 & \equiv X-Y\Delta\nu,
\end{aligned} \label{eq:r}
\]
where $\nu_{21}=1420$MHz is the 21\,cm transition rest frequency, $\nu_{0}$
a reference central frequency with corresponding redshift $z_{0}$,
and 
\[
E(z)=\sqrt{\Omega_{m}(1+z)^{3}+\Omega_{\Lambda}}.
\]
Inverting for $\nu$:
\begin{equation}
\nu=\frac{X-r}{Y}+\nu_{21}.\label{eq:nur}
\end{equation}

In the thin-shell limit, we can write:
\begin{equation}
d^2r=X^2d\Omega. 
\label{eq:thinshell}
\end{equation}
Note that in Eq.\ref{eq:thinshell} no flat-sky approximation has been made; the angular integral is still performed over the dome.

We can rewrite the delayed-transformed visibility as 
\small
\[
\begin{aligned}V(\boldsymbol{b},\tau) & =\frac{2k_{B}}{\lambda^{2}}\int\frac{d^{3}r}{X^{2}Y}A(\boldsymbol{r})T(\boldsymbol{r})\exp\left[-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\hat{\boldsymbol{r}}+\tau\right)\nu_{r}\right]\end{aligned},
\]
\normalsize
 where $d\nu=-dr/Y$ and $d^{3}r=-X^{2}Yd\Omega d\nu$. 
We have written $\nu_{r}$ as a reminder that $\nu$ and $r$ are related
by Eq. \eqref{eq:nur}.

Most existing visibility based power-spectrum pipelines for redundant drift-scan  arrays relate the power-spectrum to the conjugate square of the visibilities \citep{first-paper, paper32, Ali2015}. We would like to generalize such relations by relating the power-spectrum to the product of two visibilities from two arbitrary baselines and time-offsets. 
The beam pattern of a baseline shifts relative to the sky as the Earth rotates. Here we choose to fix the sky, and denote the rotated coordinates
of the beam pattern with the 3 dimensional rotation operator $\Gamma$:
\small
\[
\begin{aligned}V(\boldsymbol{b'},\tau') & =\frac{2k_{B}}{\lambda^{2}}\int\frac{d^{3}r}{X^{2}Y}A(\Gamma\boldsymbol{r})T(\boldsymbol{r})\exp\left[-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\Gamma\hat{\boldsymbol{r}}+\tau'\right)\nu_{r}\right]\end{aligned}.
\]
\normalsize
With implicit bounds of integrals from $-\infty$ to $\infty$, we have:
\begin{widetext}
\begin{equation}
\begin{aligned} & \langle V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau)\rangle\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\langle T^{*}(\boldsymbol{r})T(\boldsymbol{r'})\rangle\right)A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}\frac{d^{3}\kappa'}{(2\pi)^{3}}\langle T^{*}(\boldsymbol{\kappa})T(\boldsymbol{\kappa'})\rangle e^{-i(\boldsymbol{\kappa}\cdot \boldsymbol{r}-\boldsymbol{\kappa'}\cdot\boldsymbol{r'})}\right)A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}P(\kappa)e^{-i\boldsymbol{\kappa}\cdot(\boldsymbol{r}-\boldsymbol{r'})}\right)A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
 & \approx\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}rd^{3}r'\delta_{D}^{(3)}(\boldsymbol{r}-\boldsymbol{r'})A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}r|A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r})|\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\Gamma \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)\right],\\
 & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma \hat{\boldsymbol{s}},\nu)|\exp\left[-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)\right],
\end{aligned}
\label{eq:main}
\end{equation}

where in transition from cosmological coordinates back to observing coordinates we have written $\hat{\boldsymbol{r}}\equiv\hat{\boldsymbol{s}}$, and 
\begin{equation}
\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'})=\exp\left[-i\frac{2\pi}{c}\left(\boldsymbol{b}\cdot\nu_{r}\hat{\boldsymbol{r}}-\boldsymbol{b'}\cdot\nu_{r'}\Gamma\hat{\boldsymbol{r'}}\right)\right]\exp\left[-i2\pi\tau\left(\nu_{r}-\nu_{r'}\right)\right].
\end{equation}
\end{widetext}
The third equality of Eq.(\ref{eq:main}) follows from assumption of translational invariance of sky statistics, and the first inequality follows from the assumption that
the 3D power spectrum varies negligibly over the $k$-space of interest
so that $\hat{P}_{21}(k+k')\approx\hat{P}_{21}(k)$. The careful reader would have noticed that we have also assumed here a definitive relation between $k$ and $b,\tau$. In the flat-sky limit for example, such a relation would take the form $(Xk_{x},Xk_{y},Yk_{z})=\frac{2\pi}{c}(b_{x},b_{y},\tau)$. Instrumental chromaticity and foreground isolation will be examined in \ref{sec:chromaticity}. Since $\Gamma$
is a sky rotation, it does not affect $\nu$, hence we have taken $\nu_{r}$
outside the parenthesis in the phase term $\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\Gamma \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)\right]$. Notice that the phase factor $\exp\left[-i2\pi\tau\left(\nu-\nu'\right)\right]$
drops out in the end. This means that the location and magnitude of the correlation peak does not depend on delay $\tau$, an important point we shall come back to.

Since the beam pattern and bandwidth are given in $\hat{\boldsymbol{s}}$
and $\nu$, we convert the integral back to these coordinates to get
the general relation between the delay-transformed visibilities and
the power spectrum:

\begin{equation}
\begin{aligned} & \langle V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau)\rangle\\
 & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu)|\\
 & \qquad \qquad \qquad \qquad \exp\left[-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)\right].\end{aligned}
\label{eq:final}
\end{equation}

In other words the power spectrum estimate from visibilities of a baseline pair is given by 
\begin{equation}
 P(k_{b,\tau}) = \left(\frac{\lambda^{2}}{2k_{B}}\right)^{2} \frac{\langle V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau)\rangle}{\Theta}, 
 \label{eq:opp}
\end{equation}
where the weight
\begin{equation}
\Theta \equiv\int d\nu \Theta_{\nu}, 
\label{eq:Theta}
\end{equation}
and

\[
\Theta_{\nu} \equiv \int \frac{d\Omega}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu)| e^{-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)}e^{i\psi_{\nu}}.
\label{eq:Thetanu}
\]
